Choosing climate forcing for numerical simulations of the Cordilleran ice sheet

### 2013-02-13 sections/climate.tex
Added subsection 2.3 Cordilleran climates

\subsection{Cordillera climates}

@@

@@

>>>>

\subsection{Cordilleran climates}

@@

Figure~\ref{fig:temp} shows the spatial distribution of summer (JJA) air surface temperatures from the five climate datasets used in the study, and winter (DJF) air surface temperatures from the WorldClim interpolated observation data. Summer temperatures are most relevant to the glacier model as they drive summer melt.

As can be expected, temperatures decrease with latitude, and regions further inland experience colder winters. It should be noted that the temperature gradient is much stronger in the winter than it is in the summer, and that at low elevations, temperatures get well above zero during the summer months in the entire modelling domain. In other words, there is a strong seasonality contrast beween coastal and inland regions, and regions where mean annual temperatures are well below freezing do experience warm summers.

@@

Figure~\ref{fig:prec} shows the spatial distribution of winter (DJF) precipitation rates from the five climate datasets used in the study, and summer (JJA) precipitation rates from the WorldClim interpolated observation data. Winter precipitation is most relevant to the glacier model as it drives winter accumulation.

Once again expectedly, coastal regions recieve much more precipitation than inland ones due to the unborken topographical barrier of the Boundary Ranges / Coast Mountains / Cascadia. It should be noted that from a glacier mass-balance view, this contrast is made even stronger by the difference of timing of the precipitation peak through the year. Although coastal regions experience most precipitation in the accumulation season, inland regions experience dry winters and most of the precipitation falls as rain during the summer months.

In regions like Northern Yukon and Alaska, dry winters and warm summers prevent snow accumulation and glacial inception despite of mean annual temperatures being well below zero at present. In order to account for these strong gradients in seasonality, we use monthly means to drive the ice sheet model. Previous runs made with mean annual temperature and precipitation forcing and not presented here lead to unrealistic ice cover in Northern Yukon and Alaska even under present-day climate while leaving the Coast Mountains free of ice.
